---
title: "Assignment_3"
author: "Archana Gadicharla"
date: "2024-03-10"
output:
  html_document: default
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Importing the dataset.
```{r}
Univ_Bank_1 <- read.csv("C:\\Users\\archa\\Downloads\\UniversalBank.csv")
```


#Loading the required packages.
```{r}
library("caret")
```

```{r}
library("ISLR")
library("ggplot2")
library("class")
library("lattice")
library("reshape2")
```

```{r}
library("melt")
```

```{r}
#Converting into a factor variable.
```

```{r}
Univ_Bank_1$Personal.Loan <- as.factor(Univ_Bank_1$Personal.Loan)
Univ_Bank_1$Online <- as.factor(Univ_Bank_1$Online)
Univ_Bank_1$CreditCard <- as.factor(Univ_Bank_1$CreditCard)
```

```{r}
#Checking the summary of the dataset.
```

```{r}
summary(Univ_Bank_1)
```
```{r}
#Dividing the entire dataset into two parts for training and validation purposes.
#Allocating 60% of the data for training and 40% for validation.
```

```{r}
set.seed(23)
```

```{r}
Split_Index <- createDataPartition(Univ_Bank_1$Personal.Loan, p = 0.6, list = FALSE)
Train_data <- Univ_Bank_1[Split_Index,]
Validation_data <- Univ_Bank_1[-Split_Index,]
```

```{r}
dim(Train_data)
```
```{r}
dim(Validation_data)
```
```{r}
normalising1 <- preProcess(Train_data[,-c(10,14:14)],method=c("center","scale"))
Train_data1 <-predict(normalising1,Train_data)
Validation_data1 <-predict(normalising1,Validation_data)
```

A).Create a pivot table for the training data with online as a column variable,CC as a row variaable,and loan as a secondary row variable. The values inside the table should convey the count.In R use functions melt() and cast(),or function table(). In Python,use panda dataframe methods melt() and pivot().


```{r}
PTable_1 <- ftable(Train_data1[,c(14,10,13)])
PTable_1
```
```{r}
#Here 14- CreditCard, 10- Personal.Loan, 13- Online
```

B).Consider the task of classifying a customer who owns a bank credit and is actively using online banking services.
Looking at the pivot table, what is the probability that this customer will accept the loan offer?

Ans) Using the information extracted from the pivot table, we can compute the probability of the customer accepting the loan offer as 52 divided by the sum of 52 and 503, resulting in a probability of 0.096.


C) Create a two separate pivot tables for the training data. One will have Loan(rows) as a function of online(columns) and the other will have Loan(rows) as a function of CC.

```{r}
melt_1 <- melt(Train_data1,id=c("Personal.Loan"), variable="Online")
```
```{r}
melt_2 <- melt(Train_data1,id=c("Personal.Loan"), variable="CreditCard")
```

```{r}
cast_1 = dcast(melt_1,Personal.Loan~Online)
```

```{r}
cast_2 <- dcast(melt_2, Personal.Loan~CreditCard)
```
D).Compute the following quantities [P(A | B) means “the probability of A given B”]:
i. P(CC = 1 | Loan = 1) (the proportion of credit card holders among the loan acceptors)
ii. P(Online = 1 | Loan = 1)
iii. P(Loan = 1) (the proportion of loan acceptors)
iv. P(CC = 1 | Loan = 0)
v. P(Online = 1 | Loan = 0)
vi. P(Loan = 0

```{r}
ftable(Train_data1[,c(10,13)])
```
```{r}
ftable(Train_data1[,c(10,14)])
```
1. P(CC=1|Loan=1)=(92/92)=0.319
2. P(Online=1|Loan=1)=(167/167+121)=0.579
3. P(Loan=1)=(288/288+2712)=0.096
4. P(CC=1|Loan=0)=(812/812+1900)=0.299
5. P(Online=1|Loan=0)=(1624/1624+1088)=0.598
6. p(Loan=0)=(2712/2712+288)=0.904


E. Use the quantities computed above to compute the naive Bayes probability P(Loan = 1 | CC
= 1, Online = 1)?

ANS) (0.319* 0.579* 0.096*)/(0.319* 0.579*0.096)+(0.299*0.598*0.904)=0.098


F. Compare this value with the one obtained from the pivot table in (B). Which is a more
accurate estimate?

ANS) In section B, we obtained a probability value of 0.096, and in the previous question, we calculated a probability value of 0.098. Although these values show slight differences, it's crucial to acknowledge that in part B, we incorporated a more extensive range of dependent information. Hence, we can confidently assert that the value derived in part B is more precise and indicative of the underlying data.



G. Which of the entries in this table are needed for computing P(Loan = 1 | CC = 1, Online = 1)?
Run naive Bayes on the data. Examine the model output on training data, and find the entry
that corresponds to P(Loan = 1 | CC = 1, Online = 1). Compare this to the number you
obtained in (E).

```{r}
library("naivebayes")
```

```{r}
naive_b <- naive_bayes(Personal.Loan~Online+CreditCard,data=Train_data1)
naive_b
```
The customer who accepts the loan, uses a credit card, and engages in online banking yields a probability of 0.096 according to the Naive Bayes Model. This finding closely aligns with the value obtained in section E of our analysis.







