---
title: "Assignment(1)"
author: "Archana Gadicharla"
date: "2024-02-02"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r}
netflix_india <- read.csv("C:\\Users\\archa\\Downloads\\archive (2)\\netflix_india_shows_seasons.csv")
View(netflix_india)
```
This file has been imported from https://www.kaggle.com/datasets/jhajalaj/netflix-india-shows-and-movies

```{r}
mode(netflix_india$episode_count)
```
```{r}
median(netflix_india$episode_count)
```
```{r}
mean(netflix_india$episode_count)
```
```{r}
max(netflix_india$episode_count)
```
```{r}
sd(netflix_india$episode_count)
```
These values represent descriptive statistics for a selection of quantitave variables.
The above values shows mode, median, maximum,and standard deviation for the quantitave variables.

```{r}
str(netflix_india$name)
```
These values represent categorical descriptive analysis of variables.
```{r}
netflix_india_transformed <-(netflix_india$episode_count - mean(netflix_india$episode_count)/median(netflix_india$episode_count))
head(netflix_india_transformed)
```
Transformation of variables has been done above.
```{r}
hist(netflix_india$release_year)
```
The above graphical representation is a histogram.
```{r}
x <-netflix_india$release_year
y <-netflix_india$episode_count
plot(x,y, main="Scatterplot", xlab = "release_year", ylab = "episode_count" )
```
The above graphical representation is a scatterplot.
The selected variables are episode_count and release_year.



