---
title: "Goal Programming-QMM"
author: "Archana Gadicharla"
date: "2024-10-31"
output: html_document
---

problem statement:

The Research and Development Division of the Emax Corporation has developed three new
products. A decision now needs to be made on which mix of these products should be produced.
Management wants primary consideration given to three factors: total profit, stability in the
workforce, and achieving an increase in the company’s earnings next year from the $75 million
achieved this year. In particular, using the units given in the following table, they want to
Maximize Z = P - 6C - 3D, where
P = total (discounted) profit over the life of the new products,
C = change (in either direction) in the current level of employment,
D = decrease (if any) in next year’s earnings from the current year’s level.
The amount of any increase in earnings does not enter into Z, because management is concerned
primarily with just achieving some increase to keep the stockholders happy. (It has mixed
feelings about a large increase that then would be difficult to surpass in subsequent years.)
The impact of each of the new products (per unit rate of production) on each of these factors is
shown in the following table.


```{r}
library(knitr)
library(kableExtra)
```

```{r}
emaxcorpo.dtfr= data.frame(Factor=c("Total Profit","Employement level","Earning next year"),
 A=c(20,6,8),
 B=c(15,4,7),
 C=c(25,5,5),
 Goal=c("Maximize","=50",">=75"),
 Units=c("Millions of dollars",'Hundereds of employees','Million of dollars'),
 Penalties = c(" ",'(+)5,(-)5',-2)
 )
emaxcorpo.dtfr %>%
 kable(align = "c") %>%
 kable_classic() %>%
 add_header_above(header = c(" "=1,"Product"=3," "=3)) %>%
 add_header_above(header = c(" "=1,"Unit contribution"=3," "=3)) %>%
 column_spec(1,border_right = TRUE) %>%
 column_spec(4,border_right = TRUE) %>%
 column_spec(5,border_right = TRUE) %>%
 column_spec(6,border_right = TRUE)
```

Q-1: Define y1+ and y1-, respectively, as the amount over (if any) and the amount under (if any) the employment
level goal. Define y2+ and y2- in the same way for the goal regarding earnings next year. Define x1, x2, and x3 as
the production rates of Products 1, 2, and 3, respectively. With these definitions, use the goal programming
technique to express y1+, y1- , y2+ and y2- algebraically in terms of x1, x2, and x3. Also, express P in terms of x1,
x2, and x3.


###Solution.
Employment Level (Goal = 50):
The direction of the constraint for "Employment level" is “=,” suggesting that the company does not want to fall below or exceed this goal. Therefore, the penalty for this factor is bi-directional.

Earnings Next Year (Goal ≥ 75):
Since the company wants to avoid earning less than 75, the penalty for this factor is one-sided and only penalizes shortfalls, as falling below 75 would be considered harmful.


Writing Equations in Deviation Form
using the form yi= y+i + y-i
y+1 is a positive deviation or overachievement of employment level.
y-1 is a negative deviation or underachievement of employment level.
y+2 is a positive deviation or overachievement of earnings next year.
y-2 is a negative deviation or underachievement of earnings next year.

Objective Function in terms of x1, x2, x3, y+1, y-1, y+2, and  y-2
The objective function is:
 P=20x1 + 15x2 + 25x3
 
 Deviation Form of Constraints
 Employment Level: 6x1 + 4x2 + 5x3 = 50 
Writing in deviation form : 6x1 + 4x2 + 5x3 + (y-1 - y+1) = 50 

 Earnings Next Year: 8x1 + 5x2 + 7x3 >=  75
Writing in deviation form : 8x1 + 5x2 + 7x3 + (y-2 - y+2) = 75

NON-NEGATIVE CONSTRAINTS:
x1 ≥ 0, x2 ≥ 0, x3 ≥ 0, y + 1 ≥ 0, y − 1 ≥ 0, y + 2 ≥ 0, y − 2 ≥ 0


2. Express management’s objective function in terms of x1, x2, x3, y1+, y1- , y2+ and y2-?

Objective Function:
Given: 
Maximize Z = P - 6C - 3D __(1)
Substituting the Values of P, C, and D in (1)
MaxZ = 20x1 + 15x2 + 25x3 − 5(y + 1 + y − 1) − 2(y − 2)
=> MaxZ = 20x1 + 15x2 + 25x3 − 5 y + 1 − 5 y − 1 −  0 y + 2 − 2 y − 2




3. Formulate and solve the linear programming model. What are your findings?

```{r}
library(lpSolve)
library(lpSolveAPI)
```

```{r}
lp=make.lp(2,7)
set.objfn(lp,c(20,15,25,-5,-5,0,-2))
lp.control(lp,sense="max")
```


```{r}
set.row(lp,1,c(6,4,5,-1,1,0,0),indices = c(1,2,3,4,5,6,7))
set.row(lp,2,c(8,7,5,0,0,-1,1),indices = c(1,2,3,4,5,6,7))
set.rhs(lp,c(50,75))
set.constr.type(lp,c("=","="))
set.bounds(lp,lower = rep(0,7))
lp.rownames=c("Total Profit","Employement Level","Earning next year")
lp.colnames=c("x1","x2","x3","y1p","y1m","y2p")
solve(lp)
```

The value of the optimized objective function is:

```{r}
get.objective(lp)
```

The decision variables values For the optimal objective function, are:

```{r}
get.variables(lp)
```

After researching this topic, I reached the following conclusions: The Linear Programming (LP) problem was effectively solved using the provided objective function and constraints. The optimal objective value for the LP problem is 250, with the values of the decision variables in the optimal solution as follows:

x1 = 0
x2 = 0
x3 = 15
y+1 = 25
y-1 = 0
y+2 = 0
y-2 = 0



These results demonstrate an optimal solution to the goal programming problem. The values of the decision variables offer insights into the recommended employment levels, expected earnings for the coming year, and any deviations from the established targets, all within the framework of the given constraints and penalties. The objective was to reduce deviations from targets while considering the associated penalties.

In this solution, only y+1 has a non-zero value, indicating that employment level is the only metric exceeding its target. The penalty for this deviation is calculated as 5*125 . All other variables show no deviation, resulting in no additional penalties.

